---
title: "Maps"
author: "Sarah Marcrum"
format: html
editor: visual
---

### Data Source

<https://quickstats.nass.usda.gov/results/466778BB-552D-3C7C-B9B4-2BDE71F5AF1E>

## Summary

This graphic shows the primary crop in each Illinois county by acres harvested. I chose a color pallet with distinct colors to make it easy to distinguish the different crop designations for each county. Since a few counties did not have data, I chose to use a black outline on the map; this made it easy to see the county lines while also preventing the states without data (which are white) from blending into a white border. Making the map interactive allowed additional data (acres harvested, county names) to be shown in the tooltip instead of over-crowding the visualization.

```{r, message=FALSE}
#read data
data <- read.csv('IL_crops_2021.csv')
```

```{r, echo=FALSE, message=FALSE}
#filter relevant data
library(dplyr)
county <- data %>% select(County, Commodity, Data.Item, Value)
```

```{r, echo=FALSE, message=FALSE}
library(data.table)
county_harvest <- county %>% filter(Data.Item %like% "ACRES HARVESTED")

#select top crop for each county by acres harvested
top_county_harvest <- county_harvest %>% arrange(desc(Value)) %>%
  group_by(County) %>%
  slice(1)

head(top_county_harvest)
```

```{r, message=FALSE, warning=FALSE, echo=FALSE}
library(tigris)

#join map data
IL <- counties("Illinois", cb = TRUE)
IL$NAME <- toupper(IL$NAME)

IL_harvest <- IL %>% 
  left_join(top_county_harvest, by = c("NAME" = "County"))

IL_harvest <- IL_harvest %>% rename("County" = "NAME", 
                                    "Crop" = "Commodity",
                                    "Acres" = "Value")

head(IL_harvest)
```

```{r, warning=FALSE, message=FALSE}
library(ggplot2)
library(ggdendro)
library(paletteer)
library(plotly)

# define highlight key
hl <- highlight_key(IL_harvest, ~County)

# create map
p_harvest <- ggplot(data = hl) + 
  geom_sf(color="black", aes(geometry = geometry, fill=Crop, label=County, label2=Acres))
p_harvest <- p_harvest + ggdendro::theme_dendro()
p_harvest <- p_harvest + guides(fill=guide_legend(title="Top Crop by Acres Harvested"))
p_harvest <- p_harvest +  coord_sf() + scale_fill_paletteer_d("rtist::picasso") 
p_harvest <- p_harvest + ggtitle("Top Crops in Illinois Counties")

p_harvest

```

```{r, message=FALSE}
#make it interactive!

plt <- ggplotly(p_harvest, tooltip = c("County", "Crop", "Acres")) %>% highlight(
  on = "plotly_hover",
  off = "plotly_relayout"
)

plt
```
